Inducing Domain Specific Languages for Bayesian Program Learning

### 2018-02-01 paper.tex
moved stuff that is going to be in the appendix to the bottom of the document
minor edits

\noindent\textbf{Definition.} The \emph{frontier of task $x$}, written $\mathcal{F}_x$,
is a set of programs where $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.

@@

To make this concrete we need to define what programs actually are and
what form $\probability[p |\mathcal{D},\theta]$ takes.
In this work, we represent programs as \emph{polymorphicly-type $\lambda$-calculus expressions}.
$\lambda$-calculus is a formalism for expressing functional programs.
It includes variables, function application, and the ability to create new functions using ...

@@

\begin{algorithm}[tb]
   \caption{Generative model over programs}
   \label{programGenerativeModel}
   \begin{algorithmic}
     \STATE \textbf{function} sampleProgramFromDSL$(\mathcal{D}, \theta, \tau)$:
  \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, type $\tau$
  \STATE \textbf{Output:} a program whose type unifies with $\tau$
  \STATE \textbf{return} sample$(\mathcal{D}, \theta, \varnothing, \tau)$
\STATE
     \STATE \textbf{function} sample$(\mathcal{D}, \theta, \mathcal{E}, \tau)$:
  \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, environment $\mathcal{E}$, type $\tau$
  \STATE \textbf{Output:} a program whose type unifies with $\tau$
  \IF{$\tau = \alpha\to\beta$}
  \STATE var $\gets$ an unused variable name
  \STATE body $\sim$ sample$(\mathcal{D},\theta,\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$
   \STATE \textbf{return} $\lambda \text{var}.$ body
   \ENDIF
   %   \ELSE
   \STATE $\text{primitives} \gets\{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
   \STATE \hspace{2.5cm}$\text{if }\tau\text{ can unify with yield}(\tau')) \} $

@@

Why enumerate, when the program synthesis community has invented many
sophisticated algorithms that search for programs?~\cite{solar2008program,schkufza2013stochastic,feser2015synthesizing,osera2015type,polozov2015flashmeta}.
We have two reasons:% for using the simple strategy of enumerating programs:
\begin{itemize}
\item Enumeration is a general approach that can be applied to any program induction problem. Many of these more sophisticated approaches require special conditions on
  the space of of programs.
\item A key point of our work is that learning the DSL, along with a neural recognition model, can make program induction tractable, even if the
  search algorithm is very simple.  
\end{itemize}
A main drawback of an enumerative search algorithm is that we have no
efficient means of solving for arbitrary constants that might occur in the
program. In Section~\ref{regressionSection},
we will show how to find programs with real-valued constants
by automatically differentiating through the program and setting the constants using gradient descent.
In Section~\ref{textSection}
we will show that the bottom-up neural recognition model can learn
which discrete constants should be included in a program.

@@

Additionally, for each proposed $\mathcal{D}$ we have to reestimate $\theta$.
Although this problem may seem 
very similar to estimating the parameters of a probabilistic context free grammar (PCFG: see~\cite{}),
for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
$\mathcal{D}$
 is actually context-sensitive
due to the presence of variables in the programs and also due to the
polymorphic typing system.
In the Appendix we derive an efficient MAP estimator for $\theta$.

@@

\begin{figure*}
  \begin{tabular}{lll}
    \toprule
    Domain&Example programs in frontiers&Example proposed subexpression\\\midrule
    Boolean circuits&
    \begin{tabular}{l}
      \texttt{(lambda x (nand x x))}\\
      \texttt{(lambda x (lambda y (nand x (nand y y))}
    \end{tabular}&
    \texttt{(nand z z)}\\\midrule
    String editing&
    \begin{tabular}{l}
      \texttt{(lambda s (+ ',' (index 0 (split ',' s))))}\\
      \texttt{(lambda s (index 0 (split ' ' s)))}
    \end{tabular}&
    \texttt{(index 0 (split z s))}\\\bottomrule
  \end{tabular}
  \caption{The DSL induction algorithm works by proposing subexpressions of programs to add to the DSL.
    These subexpressions are taken from programs in the frontiers (middle column), and can introduce new variables (\texttt{z} in the right column).}
  \label{fragmentExample}
  \end{figure*}
\section{Program Representation}
We choose to represent programs using $\lambda$-calculus~\cite{pierce}.
A $\lambda$-calculus expression is either:
\\\noindent A \emph{primitive}, like the number 5 or
  the function \texttt{sum}.
\\\noindent A \emph{variable}, like $x$, $y$, $z$
\\\noindent A $\lambda$\emph{-abstraction}, which creates a new function. $\lambda$-abstractions have a variable and a body. The body is a $\lambda$-calculus expression. Abstractions are written as $\lambda \text{var}. \text{body}$.
\\\noindent An \emph{application} of a function to an argument. Both the function and the argument are $\lambda$-calculus expressions. The application of the function $f$ to the argument $x$ is written as $f\; x$.

@@

>>>>

\noindent\textbf{Definition.} The \emph{frontier of task $x$}, written $\mathcal{F}_x$,
is a finite set of programs where $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.

@@

To make this concrete we need to define what programs actually are and
what form $\probability[p |\mathcal{D},\theta]$ takes.
In this work, we represent programs as \emph{polymorphicly-typed $\lambda$-calculus expressions}.
$\lambda$-calculus is a formalism for expressing functional programs.
It includes variables, function application, and the ability to create new functions using ...

@@

\begin{algorithm}[tb]
   \caption{Generative model over programs}
   \label{programGenerativeModel}
   \begin{algorithmic}
     \STATE \textbf{function} sampleProgramFromDSL$(\mathcal{D}, \theta, \tau)$:
  \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, type $\tau$
  \STATE \textbf{Output:} a program whose type unifies with $\tau$
  \STATE \textbf{return} sample$(\mathcal{D}, \theta, \varnothing, \tau)$
\STATE
     \STATE \textbf{function} sample$(\mathcal{D}, \theta, \mathcal{E}, \tau)$:
  \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, environment $\mathcal{E}$, type $\tau$
  \STATE \textbf{Output:} a program whose type unifies with $\tau$
  \IF{$\tau = \alpha\to\beta$}
  \STATE var $\gets$ an unused variable name
  \STATE body $\sim$ sample$(\mathcal{D},\theta,\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$
   \STATE \textbf{return} $\lambda \text{var}.$ body
   \ENDIF
   %   \ELSE
   \STATE $\text{primitives} \gets\{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
   \STATE \hspace{2.5cm}$\text{if }\tau\text{ can unify with yield}(\tau') \} $

@@

Why enumerate, when the program synthesis community has invented many
sophisticated algorithms that search for programs?~\cite{solar2008program,schkufza2013stochastic,feser2015synthesizing,osera2015type,polozov2015flashmeta}.
We have two reasons:% for using the simple strategy of enumerating programs:
\begin{itemize}
\item A key point of our work is that learning the DSL, along with a neural recognition model, can make program induction tractable, even if the  search algorithm is very simple.  
\item Enumeration is a general approach that can be applied to any program induction problem. Many of these more sophisticated approaches require special conditions on
  the space of of programs.
\end{itemize}
A main drawback of an enumerative search algorithm is that we have no
efficient means of solving for arbitrary constants that might occur in the
program. In Section~\ref{regressionSection},
we will show how to find programs with real-valued constants
by automatically differentiating through the program and setting the constants using gradient descent.
In Section~\ref{textSection}
we will show that the bottom-up neural recognition model can learn
which discrete constants should be included in a program.

@@

Additionally, for each proposed $\mathcal{D}$ we have to reestimate $\theta$.
Although this problem may seem 
very similar to estimating the parameters of a probabilistic context free grammar (PCFG: see~\cite{}),
for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
$\mathcal{D}$
 is actually context-sensitive
due to the presence of variables in the programs and also due to the
polymorphic typing system.
In the Appendix we derive a tractable MAP estimator for $\theta$.

@@

\begin{figure*}
  \begin{tabular}{lll}
    \toprule
    Domain&Example programs in frontiers&Example proposed subexpression\\\midrule
    Boolean circuits&
    \begin{tabular}{l}
      \texttt{(lambda x (nand x x))}\\
      \texttt{(lambda x (lambda y (nand x (nand y y))}
    \end{tabular}&
    \texttt{(nand z z)}\\\midrule
    String editing&
    \begin{tabular}{l}
      \texttt{(lambda s (+ ',' (index 0 (split ',' s))))}\\
      \texttt{(lambda s (index 0 (split ' ' s)))}
    \end{tabular}&
    \texttt{(index 0 (split z s))}\\\bottomrule
  \end{tabular}
  \caption{The DSL induction algorithm works by proposing subexpressions of programs to add to the DSL.
    These subexpressions are taken from programs in the frontiers (middle column), and can introduce new variables (\texttt{z} in the right column).}
  \label{fragmentExample}
  \end{figure*}

@@

\pagebreak

\section{Program Representation}
We choose to represent programs using $\lambda$-calculus~\cite{pierce}.
A $\lambda$-calculus expression is either:
\\\noindent A \emph{primitive}, like the number 5 or
  the function \texttt{sum}.
\\\noindent A \emph{variable}, like $x$, $y$, $z$
\\\noindent A $\lambda$\emph{-abstraction}, which creates a new function. $\lambda$-abstractions have a variable and a body. The body is a $\lambda$-calculus expression. Abstractions are written as $\lambda \text{var}. \text{body}$.
\\\noindent An \emph{application} of a function to an argument. Both the function and the argument are $\lambda$-calculus expressions. The application of the function $f$ to the argument $x$ is written as $f\; x$.
